---
output:
  prettydoc::html_pretty: default
  html_document: default
---


---
title: "Mental Disorders and its relation to World Happiness Index"
author: "Saumya Seth"
date: "April, 2020"
output:
  html_document:
    df_print: paged
  prettydoc::html_pretty:
    theme: cayman
---

Steps involved-

1. Data Collection- First we collected data from various sites mentioned in bibliography.

2. Data Cleaning- Organize, clean and format data.

3. Input the data on R and code.

4. Visualization through graphs and plots.

5. Interpret the plots.

# We first examine the countries with the highest happiness index in relation to the occurrences of mental disorders prevalent there.

```{r}
library("readxl")
mhd <- read_excel("Mental health Depression disorder Data.xlsx");mhd
mhd1=mhd[-2]
s=subset(mhd1, Entity %in% c("Finland", "India","Denmark","Norway","Iceland",
                             "Netherlands","Switzerland","Sweden", 
                             "New Zealand", "Canada","Austria"))
t=subset(s, Year %in% "2017");t
summary(t)
m1=mean(t[["Schizophrenia (%)"]])
m2=mean(t[["Bipolar disorder (%)"]])
m3=mean(t[["Eating disorders (%)"]])
m4=mean(t[["Anxiety disorders (%)"]])
m5=mean(t[["Drug use disorders (%)"]])
m6=mean(t[["Depression (%)"]])
m7=mean(t[["Alcohol use disorders (%)"]])
m=c("Top 10 countries+India Mean",2017,0.2766717,0.9289069,0.5330204,5.598002,
    1.080373,3.835627,1.591701)
t1=rbind(t,m);t1
```

```{r}
austria=t1[1,c(3:9)]
a=as.numeric(austria)
m8=mean(a)

canada=t1[2,c(3:9)]
b=as.numeric(canada)
m9=mean(b)

denmark=t1[3,c(3:9)]
q=as.numeric(denmark)
m10=mean(q)

finland=t1[4,c(3:9)]
o=as.numeric(finland)
m11=mean(o)

iceland=t1[5,c(3:9)]
i=as.numeric(iceland)
m12=mean(i)

india=t1[6,c(3:9)]
d=as.numeric(india)
m13=mean(d)

netherlands=t1[7,c(3:9)]
y=as.numeric(netherlands)
m14=mean(y)

newzealand=t1[8,c(3:9)]
z=as.numeric(newzealand)
m15=mean(z)

norway=t1[9,c(3:9)]
w=as.numeric(norway)
m16=mean(w)

sweden=t1[10,c(3:9)]
k=as.numeric(sweden)
m17=mean(k)

switzerland=t1[11,c(3:9)]
v=as.numeric(switzerland)
m18=mean(v)

toptenandindiamean=t1[12,c(3:9)]
u=as.numeric(toptenandindiamean)
m19=mean(u)

mean20=c(m8,m9,m10,m11,m12,m13,m14,m15,m16,m17,m18,m19)

t2=cbind(t1,mean20);t2
```

```{r}
pie(mean20[-c(1.977757)],labels = c("Austria","Canada","Denmark","Finland","Iceland",
                                    "India","Netherlands","New Zealand","Norway",
                                    "Sweden","Switzerland"),
    main = "Average Occurances of Mental Disorders")

# The average score of the top 10 happiest countries 
# and India in terms of mental disorders is 1.977757.

sort(mean20[-m19])
# The country having the mean score of 1.351220 (India) 
# is having the least occurrences of mental disorders.

median(mean20[-m19])
# Countries having their score nearest to 1.994283 are at 
# the middle of the data set. 
# Finland's statistics are at the middle of the dataset.

kk=t2[-c(1,2)]
plot(kk)

hist(t2$mean20,xlim = c(0,4),xlab="Mean score of incidences of mental disorders in top 
     10 happiest countries and India",col="green",main="Histogram")
box()
```

# Since India itself has the least occurrences of mental disorders, let us compare it with the country having the most occurrences of mental disorders, New Zealand.

```{r}
barplot(c(1.351220,2.710621),names=c("India","New Zealand"),legend=T,ylim = c(0,4),ylab = 
          "Mean score of mental disorders",col=c("blue","red"),
        main = "Barplot")
box()
```

```{r}
whi <- read.csv("2017 Happiness Scores Data.csv")
s1=subset(whi, Country %in% c("Finland", "India","Denmark","Norway","Iceland",
                              "Netherlands","Switzerland","Sweden", "New Zealand", 
                              "Canada","Austria"));s1

# India ranks 122nd in the world happiness index while New Zealand 
# ranks eighth. This indicates that studying the mental disorder 
# data does not give an idea about the world happiness index. 

h=whi[c(1:10),c(3)];h

length(mean20)

mean21=mean20[c(-m19)]
mean22=mean21[c(-m13)]
length(mean22)
length(h)
plot(h,mean22,xlab = "Happiness score",
     ylab = "Mean of mental disorders in top 10 happiest countries", cex.lab = 0.9,
     pch = 19, col=c("green"))
cor(h,mean22)
```

# The correlation between the happiness score and mental disorder occurrences is -0.3630368. This is a negative correlation i.e. higher the happiness index, less occurrences of mental disorders. However, this is not a very strong negative correlation. This may be due to many reasons such as in countries with a lower happiness index, mental disorders may be rarely reported, there may not be enough professional/expert perspectives to classify mental disorders in such countries, stigmatization, etc.

# We now examine the countries with the lowest happiness index in relation to the occurrences of mental disorders prevalent there.

```{r}
s2=subset(mhd1, Entity %in% c("Liberia", "Burundi","Botswana",
                              "Syria","Togo","Yemen","Rawanda",
                              "Tanzania", "Guinea", "Central African Republic",
                              "South Sudan","India"))

t3=subset(s2, Year %in% "2017");t3

summary(t3)

m23=mean(t3[["Schizophrenia (%)"]])
m24=mean(t3[["Bipolar disorder (%)"]])
m25=mean(t3[["Eating disorders (%)"]])
m26=mean(t3[["Anxiety disorders (%)"]])
m27=mean(t3[["Drug use disorders (%)"]])
m28=mean(t3[["Depression (%)"]])
m29=mean(t3[["Alcohol use disorders (%)"]])
m30=c("Bottom 10 countries+India Mean",2017,0.175864,0.6648489,0.1385305,3.932043,
      0.9221215,3.674735,1.26637)
t4=rbind(t3,m30);t4
```

```{r}
burundi=t4[1,c(3:9)]
m31=mean(as.numeric(burundi))

botswana=t4[2,c(3:9)]
m32=mean(as.numeric(botswana))

car=t4[3,c(3:9)]
m33=mean(as.numeric(car))

guinea=t4[4,c(3:9)]
m34=mean(as.numeric(guinea))

india1=t4[5,c(3:9)]
m35=mean(as.numeric(india1))

liberia=t4[6,c(3:9)]
m36=mean(as.numeric(liberia))

ss=t4[7,c(3:9)]
m37=mean(as.numeric(ss))

syria=t4[8,c(3:9)]
m38=mean(as.numeric(syria))

tanzania=t4[9,c(3:9)]
m39=mean(as.numeric(tanzania))

yemen=t4[10,c(3:9)]
m40=mean(as.numeric(yemen))

togo=t4[11,c(3:9)]
m41=mean(as.numeric(togo))

bottomtenandindiamean=t4[12,c(3:9)]
m42=mean(as.numeric(bottomtenandindiamean))

mean43=c(m31,m32,m33,m34,m35,m36,m37,m38,m39,m40,m41,m42)

t5=cbind(t4,mean43);t5
```

```{r}
pie(mean43[-c(1.539216)],labels = c("Burundi","Botswana","Central African Republic","Guinea",
                                    "India","Liberia","South Sudan","Syria",
                                    "Tanzania","Yemen","Togo"),
    main = "Average Occurances of Mental Disorders")

# The average score of the bottom 10 happiest countries 
# and India in terms of mental disorders is 1.539216.

sort(mean43[-m42])
# The country having the mean score of 1.351220 (India) 
# is having the least occurrences of mental disorders.

median(mean43[-m42])
# Countries having their score nearest to 1.461732 are at 
# the middle of the data set. South Sudan's statistics are 
# at the middle of the dataset.

kk2=t5[-c(1,2)]
plot(kk2)

hist(t5$mean43,xlim = c(0,4),xlab="Mean score of incidences of mental disorders in bottom 
     10 happiest countries and India",
     col="green",main="Histogram")
box()
```

# Since India itself has the least occurrences of mental disorders, let us compare it with the country having the most occurrences of mental disorders, Yemen.

```{r}
barplot(c(1.351220,1.708384),names=c("India","Yemen"),legend=T,
        ylim = c(0,4),ylab = "Mean score of mental disorders",col=c("blue","red"),main = "Barplot")
box()
```

```{r}
whi <- read.csv("2017 Happiness Scores Data.csv")
s3=subset(whi, Country %in% c("Liberia", "Burundi","Botswana","Syria",
                              "Togo","Yemen","Rawanda","Tanzania", 
                              "Guinea", "Central African Republic","South Sudan","India"));s3

# India ranks 122nd in the world happiness index while 
# Yemen ranks 146th. 
# Even though India does actually rank above Yemen both in 
# terms of happiness and less prevalence of mental disorders, 
# we have seen that the data on mental disorders is not 
# an accurate representation while measuring the world happiness index. 

h1=whi[c(146:155),c(3)];h1

mean44=mean43[c(-m42)]
mean45=mean44[c(-m35)]

plot(h1,mean45,xlab = "Happiness score",
     ylab = "Mean of mental disorders in bottom 10 happiest countries", cex.lab = 0.8,
     pch = 19, col=c("green"))
cor(h1,mean45)
```

# The correlation between the happiness score and mental disorder occurrences is -0.5053617. This is a negative correlation i.e. lower the happiness score, higher occurrences of mental disorders. This relation was expected.

# Now, we study the correlation of Mental Health Disorders with Happiness Score of the top 10, bottom 10 countries and India.

```{r}
alldata=c("Austria","Canada","Denmark","Finland","Iceland",
"India","Netherlands","New Zealand",
"Norway","Sweden","Switzerland","Burundi",
"Botswana","Central African Republic","Guinea",
"Liberia","South Sudan","Syria",
"Tanzania","Yemen","Togo")

vv=mean43[-c(1.539216)]
vv2=vv[-c(1.35122)]

allpie=c(mean20[-c(1.977757)],vv2)
pie(allpie,labels = alldata,
main = "Average Occurances of Mental Disorders",cex=0.55)
```

```{r}
vv=c(t2$mean20,t5$mean43)
  
allhist1=hist(vv,xlim = c(0,4),xlab="Mean score of incidences of mental disorders in top and bottom
     10 happiest countries and India",col="green",
     main="Histogram of incidences of mental disorders",cex.lab=0.8)
box()
```

```{r}
pie(c(1.351220,2.710621,1.708384),labels=c("India","New Zealand",
                                          "Yemen"),
        main = "Incidences of mental health disorders in New Zealand, India and Yemen",cex.main=0.9,col=c('maroon3','royalblue1','yellow'))
```

# It is interesting to note that New Zealand which had a mean mental disorder percentage of 2.710621 is more than that of Yemen (1.708384). This is despite the fact that New Zealand is ranked 8th in the world happiness index while Yemen is ranked 146th. This, again, may be due to the reasons mentioned above.

# Let us take one of the disorders, eating disorders, and see its relationship with the World Happiness Score.

```{r}
eatingd=read.csv("Crime and Mental Health Disorders Data.csv")
all_names=c("Finland","Denmark","Norway","Iceland",
        "Netherlands","Switzerland","Sweden","New Zealand",
        "Canada","Austria","India","Haiti","Botswana",
        "Syria","Malawi","Yemen","Rwanda","Tanzania",
        "Afghanistan","Central African Republic","South Sudan")
ednumeric=as.numeric(eatingd$ED)
barplot(ednumeric,names=c("Finland","Denmark","Norway","Iceland",
        "Netherlands","Switzerland","Sweden","New Zealand",
        "Canada","Austria","India","Haiti","Botswana",
        "Syria","Malawi","Yemen","Rwanda","Tanzania",
        "Afghanistan","Central African Republic","South Sudan"),
        col=colors()[100:121],legend=T,ylim = c(0,1),
        ylab = "Mean score of eating disorders",main = "Barplot of eating disorder incidences",
        las=2,cex.names=0.7)
box()

plot(eatingd$rank,eatingd$Schiz,xlab = "Rank of country 
     (lower the rank,happier the country)",
     ylab = "Occurrences of Schizophrenia",
     pch = 19, col=c("green"),cex.lab=0.6)
plot(eatingd$rank,eatingd$Druguse,xlab = "Rank of country 
     (lower the rank,happier the country)",
     ylab = "Occurrences of drug use",
     pch = 19, col=c("red"),cex.lab=0.6)
plot(eatingd$rank,eatingd$Dep,xlab = "Rank of country 
     (lower the rank,happier the country)",
     ylab = "Occurrences of depression",
     pch = 19, col=c("orange"),cex.lab=0.6)
plot(eatingd$rank,eatingd$BPD,xlab = "Rank of country 
     (lower the rank,happier the country)",
     ylab = "Occurrences of Bipolar disorder",
     pch = 19, col=c("pink"),cex.lab=0.6)

# In some cases, the higher occurrences of mental disorders in a 
# country, the higher its happiness score.
```

```{r}
par(mfrow=c(1,2))
plot(h,mean22,xlab = "Happiness score",
     ylab = "Mean of mental disorders in top 10 happiest countries",
     pch = 19, col=c("green"),cex.lab=0.6)
cor(h,mean22)
plot(h1,mean45,xlab = "Happiness score",
     ylab = "Mean of mental disorders in bottom 10 happiest countries",
     pch = 19, col=c("green"),cex.lab=0.6)
cor(h1,mean45)
```

# As we can see, there exists  negative correlation between the happiness score and mental health disorders. This
# is as expected. However, there do exist some exceptions as can be visualy seen in the scatter plots.

# Still, as we have seen, the data on the top or bottom ten 'happiest' countries is not an accurate representation of the prevalence of mental disorders among the population in the respective countries.
---
title: "Crime and its relation to World Happiness Index"
author: "Saumya Seth"
date: "April, 2020"
output:
  prettydoc::html_pretty:
    theme: cayman
---
Steps involved-

1. Data Collection- First we collected data from various sites mentioned in bibliography.

2. Data Cleaning- Organize, clean and format data.

3. Input the data on R and code.

4. Visualization through graphs and plots.

5. Interpret the plots.

# We first examine the countries with the highest happiness index in relation to the occurrences of homicides prevalent there.

```{r}
library("readxl")
hsc <- read_excel("Crime Data.xlsx");hsc
hsc1=hsc[c(-3,-1)]
s4=subset(hsc1, Country %in% c("Finland",
                               "India","Denmark","Norway","Iceland","Netherlands",
                               "Switzerland","Sweden", "New Zealand",
                               "Canada","Austria"));s4
summary(s4)
e=as.numeric(s4$`Homicide rate`)
m46=mean(e)
m47=c("Top 10 countries+India Homicide Rate",m46)
s5=rbind(s4,m47);s5
```

```{r}
# The average homicide rate of the top 10 happiest countries 
# and India is 1.16361549145455.

sort(e)
# The country having the homicide rate of 0.4683437 (Norway) 
# is the safest country.

median(e)
# Countries having their score nearest to 0.9672286 are at 
# the middle of the data set. Austria's statistics are at the 
# middle of the dataset.

library(ggplot2)
data <- data.frame(
  name=c("Norway","Switzerland",
         "Iceland","Sweden","Netherlands",
         "Denmark","Finland","New Zealand","Canada",
         "Austria","India") ,  
  value=e
  )
ggplot(data, aes(x=name, y=value)) + 
  ggtitle("Homicide rates")+
  geom_bar(stat="identity")+
  theme(axis.text.x = element_text(angle = 90))
```

# India has the most occurrences of homicides among these countries while Norway has the least. 

```{r}
whi <- read.csv("2017 Happiness Scores Data.csv")
s6=subset(whi, Country %in% c("Finland", "India","Denmark","Norway",
                              "Iceland","Netherlands","Switzerland",
                              "Sweden", "New Zealand",
                              "Canada","Austria"));s6

# India ranks 122nd in the world happiness index while Norway 
# ranks first. This indicates that studying the crime rate 
# does indeed give an idea about the world happiness index. 
# More the crime, less safe it is and thus, less its ranking 
# in the world happiness index.

h2=whi[c(1:10),c(3)];h2

length(e)
length(h2)

e1=e[c(-m46)]
length(e1)

plot(h2,e1,xlab = "Happiness index",
     ylab = "Mean of homicides in top 10 happiest countries",pch =
       19,col=c("green"),cex.lab=0.7)
cor(h2,e1)
```

# The correlation between the happiness index and homicide occurrences is -0.5821803. This means that there exists a negative correlation between them. This is in accordance to our expectation of higher the crime in a country, lower its happiness index.

# We now examine the countries with the lowest happiness index in relation to the occurrences of homicides prevalent there.

```{r}
s7=subset(hsc1, Country %in% c("Liberia", 
                               "Burundi","Syrian Arab Republic","Togo","Yemen","Rwanda",
                               "Tanzania (United Republic of)", 
                               "Guinea", "Central African Republic","South Sudan",
                               "India"));s7

summary(s7)

e2=as.numeric(s7$`Homicide rate`)
m48=mean(e2)
m49=c("Bottom 10 countries+India Homicide Rate",m48)
s8=rbind(s7,m49);s8
```

```{r}
# The average homicide rate of the bottom 10 happiest countries 
# and India is 6.77595075390909.

sort(e2)
# The country having the homicide rate of 0.8793122 
# (Syrian Arab Republic) is the safest country in this data. 
# It is even less than Inida's homicide rate of 3.0792324419999999.

median(e2)
# Countries having their score nearest to 6.054537 are at 
# the middle of the data set. Burundi's statistics are at 
# the middle of the dataset.

data1 <- data.frame(
  name=c("India", "Syrian Arab Republic",
         "Rwanda","Tanzania (United Republic of)",
         "Togo","Liberia","Guinea", "Yemen", 
         "Burundi","South Sudan","Central African Republic") ,  
  value=e2
  )
  
ggplot(data1, aes(x=name, y=value)) + 
  ggtitle("Homicide rates")+
  geom_bar(stat="identity")+
  theme(axis.text.x = element_text(angle = 90))
```

# Central African Republic has the most occurrences of homicides among these countries while Syrian Arab Republic has the least. 

```{r}
s9=subset(whi, Country %in% c("Liberia", "Burundi",
                              "Botswana","Syria",
                              "Togo","Yemen","Rawanda",
                              "Tanzania", "Guinea", "Central African Republic",
                              "South Sudan","India"));s9

# India ranks 122nd in the world happiness index while 
# Central African Republic ranks 155th. 
# This indicates that studying the crime rate does indeed give 
# an idea about the world happiness index. 
# More the crime, less safe it is and thus, less its 
# ranking in the world happiness index.

h3=whi[c(146:155),c(3)];h3

length(e2)
length(h3)

e3=e[c(-m46)]
length(e3)

plot(h3,e3,xlab = "Happiness index",ylab = "Mean of homicides in bottom 10 happiest countries",pch = 19,col=c("green"),cex.lab=0.7)
cor(h3,e3)
```

# The correlation between the happiness index and homicide occurrences is -0.7263321. This means that there exists a strong negative correlation between them. This is in accordance to our expectation of higher the crime in a country, lower its happiness index.

```{r}
e4=c(e,e1)
library(ggplot2)
data3 <- data.frame(
  name=c("Norway","Switzerland",
         "Iceland","Sweden","Netherlands",
         "Denmark","Finland","New Zealand","Canada",
         "Austria","India","Syrian Arab Republic",
         "Rwanda","Tanzania (United Republic of)",
         "Togo","Liberia","Guinea", "Yemen", 
         "Burundi","South Sudan","Central African Republic") ,  
  value=e4
  )
ggplot(data3, aes(x=name, y=value)) + 
  ggtitle("Homicide rates")+
  geom_bar(stat="identity")+
  theme(axis.text.x = element_text(angle = 90))

par(mfrow = c(1, 2))
plot(h2,e1,xlab = "Happiness index",
     ylab = "Mean of homicides in top 10 happiest countries",pch =
       19,col=c("green"),cex.lab=0.7)
plot(h3,e3,xlab = "Happiness index",ylab = "Mean of homicides in bottom 10 happiest countries",pch = 19,col=c("green"),cex.lab=0.7)
```
